--- 
title: "HR_Sales_Bookdown"
author: "Miguel Rodicio Honrado"
date: "03-02-2026"
site: bookdown::bookdown_site
---
# Methods

In this project, we want to analyse two dataset. The first dataset we are going to use is related with the sales of a shop from September 2023 until September 2024, and the main idea is to look for the principal income of the shop, from which item we can obtain more money or which is the less bought item . The second dataset is related with the workers of data and HR-department of an industry, we can look for important resources such us Gender, Years at the Company, Job Level and the income of the worker.

## Plotting amounts

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(scales)
library(ggforce)
library(ggrepel)
library(treemapify)
```

We want to see the amount of money earned in the store with the different products offered. With this, although we would still need to analyze the price we pay to have that product in the store, we can consider which product brings us the most money and which one brings us the least. For this we first group the colums Product type and we sum the total price of that product and, after that, we separate it into the different categories for the barplot.

```{r}
sales_data <- read_csv("data/Electronic_sales_Sep2023-Sep2024.csv")

money_by_product <- sales_data %>%
  group_by(`Product Type`) %>%
  summarise(total_sales = sum(`Total Price`)) %>%
  ungroup()

ggplot(money_by_product,
       aes(x = fct_reorder(`Product Type`, total_sales),
           y = total_sales)) +
  geom_col(fill = "blue") +
  coord_flip() +
  scale_y_continuous(labels = label_comma()) +
  labs(
    title = "Total Sales of each Product Type",
    x = "Product Type",
    y = "Total Sales"
  )
```

It is also important to know which payment methods generate the most profit for us in order to identify those associated with higher purchase amounts. To do this, we calculated the mean and the median, since the mean can be affected by outliers, while the median is more robust. Thanks to this, we observed that the methods with the highest spending per purchase are PayPal and bank transfer. However, we can also see how outliers affect the mean, as it increases relatively compared to the median by about 250 per sale.

```{r}
sales_payment <- sales_data %>%
  group_by(`Payment Method`) %>%
  summarise(
    mean_price = mean(`Total Price`, na.rm = TRUE),
    median_price = median(`Total Price`, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = c(mean_price, median_price),
    names_to = "Statistic",
    values_to = "Value"
  )

ggplot(sales_payment,
       aes(x = fct_reorder(`Payment Method`, Value),
           y = Value,
           color = Statistic)) +
  geom_point(size = 5) +
  coord_flip() +
  scale_y_continuous(labels = label_comma()) +
  labs(
    title = "Mean and Median Price by Payment Method",
    x = "Payment Method",
    y = "Total Price"
  )
```